---
title: "QTL simulations"
output: 
html_document: 
keep_md: yes
---

## Intro

The goal is to determine the best design for QTL mapping in an F2 population.  Given that a certain number of plants will be grown, is it better for all of those plants to be F2s or, instead, to use a smaller number of F2s and clonally replicate them?

## Load Libraries
```{r setup}
library(qtl)
library(ggplot2)
library(reshape2)
```

## Set up parameters.
```{r parameters}
chromosomes <- 5
individuals <- 600 # number of individual plants that would be replicated.  
repl <- 3 #
```

## Simulate a genetic map.
100 markers per chromosome
```{r simmap}
map <- sim.map(len=rep(100,chromosomes),n.mar=100,include.x = FALSE)
```

## Simuate a QTL Cross
One QTL per chromosome, in the middle.  
The QTL effect size varies, increasing as we move from chromosome to chromosome.
```{r simcross}
cross <- sim.cross(map=map,
                   model = matrix(c(1:chromosomes, #QTL location (chromosome)
                                    rep(50,chromosomes), #QTL position on chromosome
                                    1:chromosomes, # additive effect sizes, ranging from 1 to # of chromosomes
                                    rep(0,chromosomes)), # dominance deviation
                                  nrow = chromosomes, byrow = FALSE), #1 QTL per chromosome at 50cM, increasing effect size 1:12
                   n.ind=repl*individuals,
                   type="f2")
cross <- sim.geno(cross)
```

## Add noise to phenotypes
Add some noise.  We create several different phenotypes, each with more noise added.
```{r}
phenotypes <- lapply(0:chromosomes, function(noise) {
  #create the plant replicates
  pheno.matrix <- t(apply(cross$pheno,1,rep,repl)) 
  
  #add the nosise
  pheno.matrix <- pheno.matrix + rnorm(n=prod(dim(pheno.matrix)), 
                                       mean=0,
                                       sd=noise)
}) #each element of the list is a phenotype matrix of 3 replicates, with increasing amounts of noise added.

#now create separate cross objects for each of the two designs
cross.no.rep <- cross

#for the no rep case take the first replicate
cross.no.rep$pheno <- as.data.frame(sapply(phenotypes, FUN = function(x) x[,1])) 

cross.with.rep <- cross

#for the case with replicates take the average of the replicates
cross.with.rep$pheno <- as.data.frame(sapply(phenotypes, rowMeans))
cross.with.rep <- cross.with.rep[,1:individuals] # so that the total number of individuals grown is the same 
```

## Do the QTL mapping
Use scanone, imputation method.
```{r map}
no.rep.scanone <- scanone(cross.no.rep,pheno.col=1:(chromosomes+1),method = "imp")
with.rep.scanone <- scanone(cross.with.rep,pheno.col=1:(chromosomes+1),method = "imp")
```

## Plot the results
```{r map2}
no.rep.scanone.melt <- melt(no.rep.scanone,id.vars = c("chr","pos"))
no.rep.scanone.melt$type <- "no replication"
with.rep.scanone.melt <- melt(with.rep.scanone,id.vars = c("chr","pos"))
with.rep.scanone.melt$type <- "with replication"
combined.scanone <- rbind(no.rep.scanone.melt,with.rep.scanone.melt)
pl <- ggplot(combined.scanone,aes(x=pos,y=value,color=variable))
pl <- pl + facet_grid(type ~ chr) 
pl <- pl + geom_line()
pl + ylab("LOD score")
```



